---
title: "HW1 - BIMS 8701"
author: "Edward Chen"
date: "2024-02-14"
output: html_document
---

# BIMS 8701 Assignment 1: Statistics Review
Due February 19, 2024
```{r}
library(Biostrings)
```
## A. This section is to help you get familiar with Bioconductor.

A.1. What version of Bioconductor are you using? (0.5 pt)
```{r}
library(BiocManager)
```
3.18
A.2. What version of BiocManager do you have installed? (0.5 pt)
1.30.22

A.3. How many total packages are available to be installed by BiocManager? Answer
this question using `BiocManager::available` (1 pt)
```{r}
length(BiocManager::available())
```
23964 

A.4. Download the Illumina TruSeq adapters bundled with the trimmomatic software
from here: https://github.com/timflutre/trimmomatic/blob/master/adapters/TruSeq3-PE-
2.fa -- then, use the Biostrings package to read the file into R, and then count the
frequency of each base for each sequence. (1 pt)

```{r}
dna_set <- readDNAStringSet("./adapters/TruSeq3-PE-2.fa")
data = letterFrequency(dna_set, letters=alphabet(dna_set))
col_sums = apply(data, 2, sum)
total_sum = sum(col_sums)
result = col_sums/total_sum
result[1:4]
```

A.5. What is the most frequently used nucleotide in these sequences overall, and what
is its overall frequency expressed as a percentage? (1 pt)

The most frequently used nucleotide is C with a overall frequency of 27.94118%.

A.6. Compute the standard deviation across sequences of the frequency of each base.
Which nucleotide has the highest standard deviation and what is it? (1 pt)
```{r}
result = apply(data, 2, sd)
result[1:4]
```
The nucleotide G has the highest standard deviation with a standard deviation of 4.335897.

## B. This section is to help you better understand the central limit theorem.

B.1. Simulate a standard normal sample with sample size n = 5, plot the sample and calculate the sample mean. (0.5 pt)
```{r}
data <- rnorm(5)
hist(data, breaks = 5)
mean(data)
```

B.2. Repeat the simulation for one million times, plot the distribution of the one million sample means. (0.5 pt)
```{r}
data <- sapply(1:1000000, function(x) mean(rnorm(5)))
hist(data, breaks = 5)
mean(data)
sd(data)
```


B.3. Now increase the sample size to 10. Repeat the above simulation and record the distribution of the one million sample means. (0.5 pt)
```{r}
data <- sapply(1:1000000, function(x) mean(rnorm(10)))
hist(data)
mean(data)
sd(data)
```

B.4. Now further increase the sample size to 20, 50, 100, and 1000. Repeat the above simulation. Show how the distribution of the one million sample means changes. (1 pt)
```{r}
for(sample_size in c(20,50,100,1000)) {
  data <- sapply(1:1000000, function(x) mean(rnorm(sample_size)))
  hist(data)
  mean(data)
  sd(data)
}
```

B.5. Now let’s change the simulation to an exponential sample. Set the rate lambda = 0.1, 0.5, and 1. Repeat B.1-B.4 and show the sample mean change under each parameter setting. (1 pt)

```{r}
data <- rexp(5)
hist(data, breaks = 5)
mean(data)
for(sample_size in c(5,10,20,50,100,1000)) {
  data <- sapply(1:1000000, function(x) mean(rexp(sample_size)))
  hist(data)
  mean(data)
  sd(data)
}
```

B.6. Now let’s change to a Poisson sample. Set lambda = 1, 5, and 10. Repeat B.1-B.4 and show the sample mean change under each parameter setting. (1 pt)
```{r}
data <- rnorm(5)
hist(data, breaks = 5)
mean(data)
for(sample_size in c(5,10,20,50,100,1000)) {
  data <- sapply(1:1000000, function(x) mean(rpois(sample_size)))
  hist(data)
  mean(data)
  sd(data)
}
```
B.7. What can we conclude from this simulation study? (0.5 pt)

Hint:
1) You can use these R functions for simulation:
rnorm()
rexp()
rpois()
2) Set an appropriate bin size when plotting a histogram.
